Sparsely Activated Networks

### 2021-09-21 ms.tex
init

\documentclass[journal]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{textgreek}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usepackage{bm}
\usepackage{bbm}
\usepackage{datatool}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usepackage{hyperref}

\newcommand\figscale{0.15}
\begin{document}
\tdplotsetmaincoords{100}{-70}
\DTLloaddb{keys_values}{tmp/keys-values.csv}

@@

\section{Introduction}
Deep Neural Networks (DNNs)~\cite{lecun2015deep} use multiple stacked layers containing weights and activation functions, that transform the input to intermediate representations during the feed-forward pass.
Using backpropagation~\cite{rumelhart1986learning} the gradient of each weight w.r.t. the error of the output is efficiently calculated and passed to an optimization function such as Stochastic Gradient Descent or Adam~\cite{kingma2014adam} which updates the weights making the output of the network converge to the desired output.
DNNs were successful in utilizing big data and powerful parallel processing units and achieved state-of-the-art performance in problems such as image~\cite{krizhevsky2012imagenet} and speech recognition~\cite{graves2013speech}.
However, these breakthroughs have come at the expense of increased description length of the learned representations, which in sparsely represented DNNs is proportional with the number of:
\begin{itemize}
	\item weights of the model and
	\item non-zero activations.
\end{itemize}

@@

Initially, bounded functions such as $sigmoid$ and $tanh$ were used, however besides producing dense activation maps they also present the vanishing gradients problem~\cite{bengio1994learning}.
Rectified Linear Units (ReLUs) were later proposed~\cite{glorot2011deep, nair2010rectified} as an activation function that solves the vanishing gradients problem and increases the sparsity of the activation maps.
Although ReLU creates exact zeros (unlike its predecessors $sigmoid$ and $tanh$), its activation map consists of sparsely separated but still dense areas (Fig.~\ref{fig:activationfunctions}\subref{subfig:relu}) instead of sparse spikes.
The same applies for generalizations of ReLU, such as Parametric ReLU~\cite{he2015delving} and Maxout~\cite{goodfellow2013maxout}.
Recently, in $k$-Sparse Autoencoders~\cite{makhzani2013k} the authors used an activation function that applies thresholding until the $k$ most active activations remain, however this non-linearity covers a limited area of the activation map by creating sparsely disconnected dense areas (Fig.~\ref{fig:activationfunctions}\subref{subfig:topkabsolutes}), similar to the ReLU case.

\begin{figure*}[!t]
	\centering
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-identity-2-activations-0}}
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-relu-2-activations-0}}
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-top-k-absolutes-2-activations-0}}
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-extrema-pool-idx-2-activations-0}}
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-extrema-2-activations-0}}
	\\
	\setcounter{subfigure}{0}
	\subfloat[Identity]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-identity-2-activations-1}\label{subfig:identity}}
	\subfloat[ReLU]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-relu-2-activations-0}\label{subfig:relu}}
	\subfloat[top-k absolutes]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-top-k-absolutes-2-activations-0}\label{subfig:topkabsolutes}}
	\subfloat[Extrema-Pool indices]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-extrema-pool-idx-2-activations-0}\label{subfig:extremapoolindices}}
	\subfloat[Extrema]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-extrema-2-activations-0}\label{subfig:extrema}}
	\caption{Visualization of the activation maps of five activation functions (Identity, ReLU, top-k absolutes, Extrema-Pool indices and Extrema) for 1D and 2D input in the top and bottom row respectively.
	The 1D input to the activation functions is denoted with the continuous transparent green line using an example from the UCI dataset.
	The output of each activation function is denoted with the cyan stem lines with blue markers.
	The 2D example depicts only the output of the activation functions using an example from the MNIST dataset.
	}
	\label{fig:activationfunctions}
\end{figure*}

@@

In section~\ref{sec:flithos} we define the $\varphi$ metric, then in section~\ref{sec:sans} we define the five tested activation functions along with the architecture and training procedure of SANs, in section~\ref{sec:experiments} we experiment SANs on the Physionet~\cite{goldberger2000physiobank}, UCI-epilepsy~\cite{andrzejak2001indications}, MNIST~\cite{lecun1998gradient} and FMNIST~\cite{xiao2017fashion} databases and provide visualizations of the intermediate representations and results.
In section~\ref{sec:discussion} we discuss the findings of the experiments and the limitations of SANs and finally in section~\ref{sec:conclusions} we present the concluding remarks and the future work.

\section{$\varphi$ Metric}
\label{sec:flithos}
Let $M$ be a model with $q$ kernels each one with $m^{(i)}$ samples and a reconstruction loss function $\mathcal{L}$ for which we have:
\begin{equation}
	\label{eq:model}
	M: \bm{x} \longmapsto \hat{\bm{x}}
\end{equation}

@@

This definition of $\tilde{\mathcal{L}}$ allows us to set a base of reference for $\tilde{\mathcal{L}}$ in cases when the reconstruction the model performs is equivalent with a model that performs constant reconstructions independent of the input.
Finally using Equations~\ref{eq:compressionratio} and \ref{eq:normalizedreconstrucionloss} we define the $\varphi$\footnote{The use of the symbol $\varphi$ comes from the first character of the combined greek word {\textgreekfont φλύθος} = {\textgreekfont φλύ+θος} (flithos: pronounced as flee-thos). It consists of the first part of the word {\textgreekfont φλύ-αρος} (meaning verbose) and the second part of {\textgreekfont λά-θος} (meaning wrong). {\textgreekfont Φλύθος} is literally defined as: \textit{Giving inaccurate information using many words; the state of being wrong and wordy at the same time}.} metric of $\bm{x}$ w.r.t $M$ as follows:
\begin{equation}
	\label{eq:flithos}
	\varphi = \lVert(CR^{-1}, \tilde{\mathcal{L}}(\hat{\bm{x}},\bm{x}))\rVert_2
\end{equation}

@@

\section{Sparsely Activated Networks}
\label{sec:sans}

\subsection{Sparse Activation Functions}
\label{sec:safs}
In this subsection we define five activation functions $\phi$ and their corresponding sparsity density parameter $d^{(i)}$ for which we have:
\begin{equation}
	\label{eq:phi}
	\phi: s \longmapsto \alpha
\end{equation}

@@

\subsubsection{Identity}
\label{sec:identity}
$\phi = \mathbbm{1}$.
The Identity activation function serves as a baseline and does not change its input as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:identity}.
For this case $d^{(i)}$ does not apply.

\subsubsection{ReLU}
\label{sec:relu}
$\phi = ReLU(s)$.
The ReLU activation function produces sparsely disconnected but internally dense areas as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:relu} instead of sparse spikes.
For this case $d^{(i)}$ does not apply.

\subsubsection{top-k absolutes}
\label{sec:topkabsolutes}
The top-k absolutes function (defined at Algorithm~\ref{alg:topkabsolutes}) keeps the indices of $k$ activations with the largest absolute value and zeros out the rest, where $1 \le k < n \in \mathbb{N}$.
We set $d^{(i)} = k$, where $k = \lfloor n/m \rfloor^{\dim(\bm{x})}$.
Top-k absolutes is sparser than ReLU but some extrema are overactivated related to some others that are not activated at all, as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:topkabsolutes}.

\begin{algorithm}[H]
	\caption{top-k absolutes}
	\label{alg:topkabsolutes}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE $s$, $k$
		\ENSURE $\alpha$
		\STATE $\alpha(i) \leftarrow 0, i=1\ldots card(s)$
		\STATE $e \leftarrow topk(\lvert s\rvert, k)$
		\\\textit{\scriptsize \# e denotes the vector with the extrema indices}
		\FOR {$i$ = 0 to $card(s)$}
		\STATE $\alpha(e(i)) \leftarrow s(e(i))$
		\ENDFOR
		\RETURN $\alpha$
	\end{algorithmic}
\end{algorithm}

\subsubsection{Extrema-Pool indices}
\label{sec:extremapoolindices}
The Extrema-Pool indices activation function (defined at Algorithm~\ref{alg:extremapoolindices}) keeps only the index of the activation with the maximum absolute amplitude from each region outlined by a grid as granular as the kernel size $m^{(i)}$ and zeros out the rest.
It consists of a max-pooling layer followed by a max-unpooling layer with the same parameters while the sparsity parameter $d^{(i)}$ in this case is set $d^{(i)} = m^{(i)} < n \in \mathbb{N}$.
This activation function creates sparser activation maps than top-k absolutes however in cases where the pool grid is near a peak or valley this region is activated twice (as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:extremapoolindices}).

\begin{algorithm}[H]
	\caption{Extrema-Pool indices}
	\label{alg:extremapoolindices}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE $s$, $m$
		\ENSURE $\alpha$
		\STATE $e \leftarrow maxpool(\lvert s\rvert, m)$
		\STATE $\alpha \leftarrow maxunpool(s(e), e, m)$
		\RETURN $\alpha$
	\end{algorithmic}
\end{algorithm}

\subsubsection{Extrema}
\label{sec:extrema}
The Extrema activation function (defined at Algorithm~\ref{alg:extrema}) detects candidate extrema using zero crossing of the first derivative, then sorts them in an descending order and gradually eliminates those extrema that have less absolute amplitude than a neighboring extrema within a predefined minimum extrema distance ($med$).
Imposing a $med$ on the extrema detection algorithm makes $\bm{\alpha}$ sparser than the previous cases and solves the problem of double extrema activations that Extrema-Pool indices have (as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:extrema}).
The sparsity parameter in this case is set $d^{(i)} = med$, where $1 \le med < n \in \mathbb{N}$ is the minimum extrema distance.
We set $med = m^{(i)}$ for utilizing fair comparison between the sparse activation functions.
Specifically for Extrema activation function we introduce a `border tolerance' parameter to allow neuron activation within another neuron activated area.

\begin{algorithm}[H]
	\caption{Extrema detection with minimum extrema distance $med$}
	\label{alg:extrema}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE $s$, $med$
		\ENSURE $\alpha$
		\STATE $peaks \leftarrow \left(\frac{d s}{d t}^+ \geq 0\right) \land \left(\frac{d s}{d t}^- < 0\right)$
		\STATE $valleys \leftarrow \left(\frac{d s}{d t}^+ < 0\right) \land \left(\frac{d s}{d t}^- \geq 0\right)$
		\\\textit{\scriptsize \# + and - denote one sample padding to the right and left respectively}
		\STATE $z \leftarrow peaks \lor valleys$
		\STATE $e \leftarrow z > 0$
		\STATE $e_{ind} \leftarrow sort(z)$
		\STATE $e_{sorted} \leftarrow e(e_{ind})$
		\STATE $e_{sec}(i) \leftarrow 0, i=1\ldots card(s)$
		\\\textit{\scriptsize \# $e_{sec}$ denotes a boolean vector for the secondary extrema indices}
		\FOR {$i$ = 0 to $card(s)$}
		\IF {$\lnot e_{sec}(i)$}
		\STATE $e_r \leftarrow e \geq e_{ind}(i) - med$
		\STATE $e_l \leftarrow e \leq e_{ind}(i) + med$
		\STATE $e_m \leftarrow e_r \land e_l$
		\STATE $e_{sec} \leftarrow e_{sec} \lor e_m$
		\STATE $e_{sec}(i) \leftarrow 0$
		\ENDIF
		\ENDFOR
		\STATE $\alpha_{ind} \leftarrow e_{sorted}(\lnot e_{sec})$
		\\\textit{\scriptsize \# $\alpha_{ind}$ denotes the indices of $\alpha$ with non-zero value}
		\STATE $\alpha(i) \leftarrow 0, i=1\ldots card(s)$
		\FOR {$i$ = 0 to $card(s)$}
		\STATE $\alpha(\alpha_{ind}(i)) \leftarrow s(\alpha_{ind}(i))$
		\ENDFOR
		\RETURN $\alpha$
	\end{algorithmic}
\end{algorithm}

@@

			\node[opacity=0] at (0, 0)(input){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-signal}};

@@

			\node[opacity=0, right=0.5cm of loss] (reconstructed){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-reconstructed}};

@@

			\node[opacity=0.8, below=0.5cm of plus] (reconstruction){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-reconstruction-0}};

@@

			\node[opacity=0.8, below=0.5cm of conv2] (extrema){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-activations-0}};

@@

			\node[opacity=0.8, left=0.55cm of phi] (similarity){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-similarity-0}};

@@

			\node[left=1.5cm of conv2] (kernel){\includegraphics[scale=0.1]{tmp/UCI-epilepsy-extrema-2-kernel-0}};

@@

			\node[] at (0, 0)(input){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-signal}};

@@

			\node[right=0.5cm of loss] (reconstructed){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-reconstructed}};

@@

			\node[opacity=0, below=0.5cm of plus] (reconstruction){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-reconstruction-0}};

@@

			\node[opacity=0, below=0.5cm of conv2] (extrema){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-activations-0}};

@@

			\node[opacity=0, left=0.55cm of phi] (similarity){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-similarity-0}};
			\node[opacity=0, left=1.5cm of conv2] (kernel){\includegraphics[scale=0.1]{tmp/UCI-epilepsy-extrema-2-kernel-0}};

@@

			\node[opacity=0.8, below=0.5cm of plus] (reconstruction){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-reconstruction-1}};

@@

			\node[opacity=0.8, below=0.5cm of conv2] (extrema){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-activations-1}};

@@

			\node[opacity=0.8, left=0.55cm of phi] (similarity){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-2-similarity-1}};

@@

			\node[left=1.5cm of conv2] (kernel){\includegraphics[scale=0.1]{tmp/UCI-epilepsy-extrema-2-kernel-1}};

@@

			\node[opacity=0] at (0, 0)(input){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2-signal}};

@@

			\node[opacity=0, right=0.5cm of loss] (reconstructed){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2-reconstructed}};

@@

			\node[opacity=0.8, below=0.5cm of plus] (reconstruction){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2-reconstruction-0}};

@@

			\node[opacity=0.8, below=0.5cm of conv2] (extrema){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2-activations-0}};

@@

			\node[opacity=0.8, left=0.55cm of phi] (similarity){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2-similarity-0}};
			\node[opacity=0.8, left=1.2cm of conv2] (kernel){\includegraphics[scale=0.1]{tmp/MNIST-extrema-2-kernel-0}};

@@

			\node[] at (0, 0)(input){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2-signal}};

@@

			\node[right=0.5cm of loss] (reconstructed){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2-reconstructed}};

@@

	\label{fig:sans}

>>>>

\documentclass[journal]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{textgreek}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usepackage{bm}
\usepackage{dsfont}
\usepackage{datatool}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algorithmic}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usepackage{hyperref}

\newcommand\figscale{0.15}
\begin{document}
\tdplotsetmaincoords{100}{-70}
\DTLloaddb{keys-values}{tmp/keys-values.csv}

@@

\section{Introduction}
Deep Neural Networks (DNNs)~\cite{lecun2015deep} use multiple stacked layers containing weights and activation functions, that transform the input to intermediate representations during the feed-forward pass.
Using backpropagation~\cite{rumelhart1986learning} the gradient of each weight w.r.t.\ the error of the output is efficiently calculated and passed to an optimization function such as Stochastic Gradient Descent or Adam~\cite{kingma2014adam} which updates the weights making the output of the network converge to the desired output.
DNNs were successful in utilizing big data and powerful parallel processing units and achieved state-of-the-art performance in problems such as image~\cite{krizhevsky2012imagenet} and speech recognition~\cite{graves2013speech}.
However, these breakthroughs have come at the expense of increased description length of the learned representations, which in sparsely represented DNNs is proportional with the number of:
\begin{itemize}
	\item weights of the model and
	\item non-zero activations.
\end{itemize}

@@

Initially, bounded functions such as $sigmoid$ and $\tanh$ were used, however besides producing dense activation maps they also present the vanishing gradients problem~\cite{bengio1994learning}.
Rectified Linear Units (ReLUs) were later proposed~\cite{glorot2011deep, nair2010rectified} as an activation function that solves the vanishing gradients problem and increases the sparsity of the activation maps.
Although ReLU creates exact zeros (unlike its predecessors $sigmoid$ and $\tanh$), its activation map consists of sparsely separated but still dense areas (Fig.~\ref{fig:activationfunctions}\subref{subfig:relu}) instead of sparse spikes.
The same applies for generalizations of ReLU, such as Parametric ReLU~\cite{he2015delving} and Maxout~\cite{goodfellow2013maxout}.
Recently, in $k$-Sparse Autoencoders~\cite{makhzani2013k} the authors used an activation function that applies thresholding until the $k$ most active activations remain, however this non-linearity covers a limited area of the activation map by creating sparsely disconnected dense areas (Fig.~\ref{fig:activationfunctions}\subref{subfig:topkabsolutes}), similar to the ReLU case.

\begin{figure*}[!t]
	\centering
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-identity-1d-2-activations-0.pdf}}
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-relu-1d-2-activations-0.pdf}}
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-top-k-absolutes-1d-2-activations-0.pdf}}
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-extrema-pool-indices-1d-2-activations-0.pdf}}
	\subfloat{\includegraphics[width=0.2\textwidth]{tmp/UCI-epilepsy-extrema-1d-2-activations-0.pdf}}
	\\
	\setcounter{subfigure}{0}
	\subfloat[Identity]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-identity-2d-2-activations-1.pdf}\label{subfig:identity}}
	\subfloat[ReLU]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-relu-2d-2-activations-0.pdf}\label{subfig:relu}}
	\subfloat[top-k absolutes]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-top-k-absolutes-2d-2-activations-0.pdf}\label{subfig:topkabsolutes}}
	\subfloat[Extrema-Pool indices]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-extrema-pool-indices-2d-2-activations-0.pdf}\label{subfig:extremapoolindices}}
	\subfloat[Extrema]{\includegraphics[width=0.2\textwidth]{tmp/MNIST-extrema-2d-2-activations-0.pdf}\label{subfig:extrema}}
	\caption{Visualization of the activation maps of five activation functions (Identity, ReLU, top-k absolutes, Extrema-Pool indices and Extrema) for 1D and 2D input in the top and bottom row respectively.
	The 1D input to the activation functions is denoted with the continuous transparent green line using an example from the UCI dataset.
	The output of each activation function is denoted with the cyan stem lines with blue markers.
	The 2D example depicts only the output of the activation functions using an example from the MNIST dataset.
	}\label{fig:activationfunctions}
\end{figure*}

@@

In Section~\ref{sec:flithos} we define the $\varphi$ metric, then in Section~\ref{sec:sans} we define the five tested activation functions along with the architecture and training procedure of SANs, in Section~\ref{sec:experiments} we experiment SANs on the Physionet~\cite{goldberger2000physiobank}, UCI-epilepsy~\cite{andrzejak2001indications}, MNIST~\cite{lecun1998gradient} and FMNIST~\cite{xiao2017fashion} databases and provide visualizations of the intermediate representations and results.
In Section~\ref{sec:discussion} we discuss the findings of the experiments and the limitations of SANs and finally in Section~\ref{sec:conclusions} we present the concluding remarks and the future work.

\section{$\varphi$ Metric}\label{sec:flithos}
Let $M$ be a model with $q$ kernels each one with $m^{(i)}$ samples and a reconstruction loss function $\mathcal{L}$ for which we have:
\begin{equation}
	\label{eq:model}
	M: \bm{x} \longmapsto \hat{\bm{x}}
\end{equation}

@@

This definition of $\tilde{\mathcal{L}}$ allows us to set a base of reference for $\tilde{\mathcal{L}}$ in cases when the reconstruction the model performs is equivalent with a model that performs constant reconstructions independent of the input.
Finally using Equations~\ref{eq:compressionratio} and~\ref{eq:normalizedreconstrucionloss} we define the $\varphi$\footnote{The use of the symbol $\varphi$ comes from the first character of the combined greek word {\textgreekfont{φλύθος}} = {\textgreekfont{φλύ+θος}} (flithos: pronounced as flee-thos). It consists of the first part of the word {\textgreekfont{φλύ-αρος}} (meaning verbose) and the second part of {\textgreekfont{λά-θος}} (meaning wrong). {\textgreekfont{Φλύθος}} is literally defined as: \textit{Giving inaccurate information using many words; the state of being wrong and wordy at the same time}.} metric of $\bm{x}$ w.r.t $M$ as follows:
\begin{equation}
	\label{eq:flithos}
	\varphi=\lVert(CR^{-1},\tilde{\mathcal{L}}(\hat{\bm{x}},\bm{x}))\rVert_2
\end{equation}

@@

\section{Sparsely Activated Networks}\label{sec:sans}

\subsection{Sparse Activation Functions}\label{sec:safs}
In this subsection we define five activation functions $\phi$ and their corresponding sparsity density parameter $d^{(i)}$ for which we have:
\begin{equation}
	\label{eq:phi}
	\phi: s \longmapsto \alpha
\end{equation}

@@

\subsubsection{Identity}\label{sec:identity}
$\phi = \mathds{1}$.
The Identity activation function serves as a baseline and does not change its input as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:identity}.
For this case $d^{(i)}$ does not apply.

\subsubsection{ReLU}\label{sec:relu}
$\phi = ReLU(s)$.
The ReLU activation function produces sparsely disconnected but internally dense areas as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:relu} instead of sparse spikes.
For this case $d^{(i)}$ does not apply.

\subsubsection{top-k absolutes}\label{sec:topkabsolutes}
The top-k absolutes function (defined at Algorithm~\ref{alg:topkabsolutes}) keeps the indices of $k$ activations with the largest absolute value and zeros out the rest, where $1 \le k < n \in \mathbb{N}$.
We set $d^{(i)} = k$, where $k = \lfloor n/m \rfloor^{\dim(\bm{x})}$.
Top-k absolutes is sparser than ReLU but some extrema are overactivated related to some others that are not activated at all, as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:topkabsolutes}.

\begin{algorithm}[H]
	\caption{top-k absolutes}\label{alg:topkabsolutes}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE{$s$, $k$}
		\ENSURE{$\alpha$}
		\STATE{$\alpha(i) \leftarrow 0, i=1\ldots card(s)$}
		\STATE{$e \leftarrow topk(\lvert s\rvert, k)$}
		\\\textit{\scriptsize \# e denotes the vector with the extrema indices}
		\FOR{$i$ = 0 to $card(s)$}
		\STATE{$\alpha(e(i)) \leftarrow s(e(i))$}
		\ENDFOR{}
		\RETURN{$\alpha$}
	\end{algorithmic}
\end{algorithm}

\subsubsection{Extrema-Pool indices}\label{sec:extremapoolindices}
The Extrema-Pool indices activation function (defined at Algorithm~\ref{alg:extremapoolindices}) keeps only the index of the activation with the maximum absolute amplitude from each region outlined by a grid as granular as the kernel size $m^{(i)}$ and zeros out the rest.
It consists of a max-pooling layer followed by a max-unpooling layer with the same parameters while the sparsity parameter $d^{(i)}$ in this case is set $d^{(i)} = m^{(i)} < n \in \mathbb{N}$.
This activation function creates sparser activation maps than top-k absolutes however in cases where the pool grid is near a peak or valley this region is activated twice (as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:extremapoolindices}).

\begin{algorithm}[H]
	\caption{Extrema-Pool indices}\label{alg:extremapoolindices}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE{$s$, $m$}
		\ENSURE{$\alpha$}
		\STATE{$e \leftarrow maxpool(\lvert s\rvert, m)$}
		\STATE{$\alpha \leftarrow maxunpool(s(e), e, m)$}
		\RETURN{$\alpha$}
	\end{algorithmic}
\end{algorithm}

\subsubsection{Extrema}\label{sec:extrema}
The Extrema activation function (defined at Algorithm~\ref{alg:extrema}) detects candidate extrema using zero crossing of the first derivative, then sorts them in an descending order and gradually eliminates those extrema that have less absolute amplitude than a neighboring extrema within a predefined minimum extrema distance ($med$).
Imposing a $med$ on the extrema detection algorithm makes $\bm{\alpha}$ sparser than the previous cases and solves the problem of double extrema activations that Extrema-Pool indices have (as shown in Fig.~\ref{fig:activationfunctions}\subref{subfig:extrema}).
The sparsity parameter in this case is set $d^{(i)} = med$, where $1 \le med < n \in \mathbb{N}$ is the minimum extrema distance.
We set $med = m^{(i)}$ for utilizing fair comparison between the sparse activation functions.
Specifically for Extrema activation function we introduce a `border tolerance' parameter to allow neuron activation within another neuron activated area.

\begin{algorithm}[H]
	\caption{Extrema detection with minimum extrema distance $med$}\label{alg:extrema}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}}
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\REQUIRE{$s$, $med$}
		\ENSURE{$\alpha$}
		\STATE{$peaks \leftarrow \left(\frac{d s}{d t}^+ \geq 0\right) \land \left(\frac{d s}{d t}^- < 0\right)$}
		\STATE{$valleys \leftarrow \left(\frac{d s}{d t}^+ < 0\right) \land \left(\frac{d s}{d t}^- \geq 0\right)$}
		\\\textit{\scriptsize \# + and - denote one sample padding to the right and left respectively}
		\STATE{$z \leftarrow peaks \lor valleys$}
		\STATE{$e \leftarrow z > 0$}
		\STATE{$e_{ind} \leftarrow sort(z)$}
		\STATE{$e_{sorted} \leftarrow e(e_{ind})$}
		\STATE{$e_{seco}(i) \leftarrow 0, i=1\ldots card(s)$}
		\\\textit{\scriptsize \# $e_{seco}$ denotes a boolean vector for the secondary extrema indices}
		\FOR{$i$ = 0 to $card(s)$}
		\IF{$\lnot e_{seco}(i)$}
		\STATE{$e_r \leftarrow e \geq e_{ind}(i) - med$}
		\STATE{$e_l \leftarrow e \leq e_{ind}(i) + med$}
		\STATE{$e_m \leftarrow e_r \land e_l$}
		\STATE{$e_{seco} \leftarrow e_{seco} \lor e_m$}
		\STATE{$e_{seco}(i) \leftarrow 0$}
		\ENDIF{}
		\ENDFOR{}
		\STATE{$\alpha_{ind} \leftarrow e_{sorted}(\lnot e_{seco})$}
		\\\textit{\scriptsize \# $\alpha_{ind}$ denotes the indices of $\alpha$ with non-zero value}
		\STATE{$\alpha(i) \leftarrow 0, i=1\ldots card(s)$}
		\FOR{$i$ = 0 to $card(s)$}
		\STATE{$\alpha(\alpha_{ind}(i)) \leftarrow s(\alpha_{ind}(i))$}
		\ENDFOR{}
		\RETURN{$\alpha$}
	\end{algorithmic}
\end{algorithm}

@@

			\node[opacity=0] at (0, 0)(input){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-signal.pdf}};

@@

			\node[opacity=0, right=0.5cm of loss] (reconstructed){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-reconstructed.pdf}};

@@

			\node[opacity=0.8, below=0.5cm of plus] (reconstruction){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-reconstruction-0.pdf}};

@@

			\node[opacity=0.8, below=0.5cm of conv2] (extrema){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-activations-0.pdf}};

@@

			\node[opacity=0.8, left=0.55cm of phi] (similarity){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-similarity-0.pdf}};

@@

			\node[left=1.5cm of conv2] (kernel){\includegraphics[scale=0.1]{tmp/UCI-epilepsy-extrema-1d-2-kernel-0.pdf}};

@@

			\node[] at (0, 0)(input){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-signal.pdf}};

@@

			\node[right=0.5cm of loss] (reconstructed){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-reconstructed.pdf}};

@@

			\node[opacity=0, below=0.5cm of plus] (reconstruction){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-reconstruction-0.pdf}};

@@

			\node[opacity=0, below=0.5cm of conv2] (extrema){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-activations-0.pdf}};

@@

			\node[opacity=0, left=0.55cm of phi] (similarity){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-similarity-0.pdf}};
			\node[opacity=0, left=1.5cm of conv2] (kernel){\includegraphics[scale=0.1]{tmp/UCI-epilepsy-extrema-1d-2-kernel-0.pdf}};

@@

			\node[opacity=0.8, below=0.5cm of plus] (reconstruction){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-reconstruction-1.pdf}};

@@

			\node[opacity=0.8, below=0.5cm of conv2] (extrema){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-activations-1.pdf}};

@@

			\node[opacity=0.8, left=0.55cm of phi] (similarity){\includegraphics[scale=\figscale]{tmp/UCI-epilepsy-extrema-1d-2-similarity-1.pdf}};

@@

			\node[left=1.5cm of conv2] (kernel){\includegraphics[scale=0.1]{tmp/UCI-epilepsy-extrema-1d-2-kernel-1.pdf}};

@@

			\node[opacity=0] at (0, 0)(input){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2d-2-signal.pdf}};

@@

			\node[opacity=0, right=0.5cm of loss] (reconstructed){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2d-2-reconstructed.pdf}};

@@

			\node[opacity=0.8, below=0.5cm of plus] (reconstruction){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2d-2-reconstruction-0.pdf}};

@@

			\node[opacity=0.8, below=0.5cm of conv2] (extrema){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2d-2-activations-0.pdf}};

@@

			\node[opacity=0.8, left=0.55cm of phi] (similarity){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2d-2-similarity-0.pdf}};
			\node[opacity=0.8, left=1.2cm of conv2] (kernel){\includegraphics[scale=0.1]{tmp/MNIST-extrema-2d-2-kernel-0.pdf}};

@@

			\node[] at (0, 0)(input){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2d-2-signal.pdf}};

@@

			\node[right=0.5cm of loss] (reconstructed){\includegraphics[scale=\figscale]{tmp/MNIST-extrema-2d-2-reconstructed.pdf}};

@@

	}\label{fig:sans}
